Post-Disturbance Tree Community Trajectories in a Neotropical Forest

### 2018-04-16 WholePlotTrajectories.Rmd
Discussion

```{r Fig5, out.width = '100%', fig.asp = 1.5/3, echo=FALSE,fig.cap="Trajectories of the difference to reference inventory over 30 years after disturbance of plots communities **(a)** Richness, **(b)** Shannon and **(c)** Simpson diversities. Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation. Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3.", fig.env="figure*"}
load("DB/ReplacementTraj_ForGraphs")

@@

## Validation of the IDH

The enhanced communities evenness and functional richness after disturbance confirmed the IDH for all treatment intensity.
The divergence from the initial state however was very variable among plots, specifically after intense disturbance, which meant a strong dependency of plots recovery to initial composition or to abiotic conditions.
All plots, though, followed the same unimodal trajectory which confirmed both the resilience of forests taxonomic and functional composition and the convergence of communities towards initial states [@Hubbell1999;@Molino2001;@Baraloto2012a].
Functional traits are the most direct link between biodiversity and ecosystem functioning [@Diaz2005] the recovery of communities functional characteristics would go along the recovery of ecosystems processes although taxonomic composition remain altered [@Guariguata2001].
Disturbance proved to favor species with acquisitive strategies, displaying efficient resource acquisition (high SLA) but short-lived tissues (low wood density, leaf toughness and thickness) and with a dispresive strategy (low seed mass) [@Westoby1998;@Wright2004;@Reich2014].
#The comparison of observed trajectories to a null model with random traits set shuffling showed an oriented functional shift after disturbance which suggest additional biotic and abiotic filters for the 10 to 15 years after disturbance.
#This means a shift from exclusive competition favoring competitive species before disturbance to environmental filtering selecting acquisitive and disturbance resistant species.
In particular, the evolution seed mass classes proportions revealed a clear increase in small seeded, therefore large dispersive, species [@TerSteege2001;@Flores2006;@Haddad2008].

>>>>

```{r Fig5, out.width = '100%', fig.asp = 1.5/3, echo=FALSE,fig.cap="Trajectories of the difference to reference inventory over 30 years after disturbance of plots communities **(a)** Richness, **(b)** Shannon and **(c)** Simpson diversities. Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation. Initial treatments are represented by solid lines colorswith green for control, blue for T1,orange for T2 and red for T3.", fig.env="figure*"}
load("DB/ReplacementTraj_ForGraphsN")

@@

## A validation of the intermediate disturbance hypothesis

The monitoring 30 years of disturbed forest communities confirmed the limited direct impact of low disturbance on species richness as it was found for impact survey of selective logging (@Cannon1998;@Baraloto2010).
Only the most disturbed plots had not fully achieved the ongoing recovery of genus richness, although all had already reached equivalent levels as those of some control plots.
Such long time richness recovery had already been observed in several logging experiments, which also highlighted the role of random disturbance damage in the richness decrease (@deAvila2015;@Hu2018).

The evenness of all disturbed plots and the richness of low disturbance plots on the other hand mainly followed an asymptotic growth with a sharp increase until 15 years after disturbance.
Such increased evenness is explained by a much more homogeneous species distribution after disturbance.
Along with the IDH, disturbance then enhanced the growth and survival of previously infrequent species, reorganizing the typical high dominance structure of hyperdiverse mature forests.
This increase in taxonomic diversity was accompanied by an increased taxonomic dissimilarity to initial state along time, translating a a taxonomic turnover benefiting to pioneers and light demanding species, as illustrated by the communities functional shifts towards resource-acquisitive strategies (sharp increase in the SLA, leaf thickness and bark thickness and decrease in wood density, leaf toughness and maximum height) [@Westoby1998;@Wright2004;@Reich2014].
In line with the IDH, the post-disturbance changes in abiotic environment and competitive pressure then well favored pioneers outcompeting other species when abiotic resources are abundant but otherwise excluded in mature forests by long-lived, resistant and shade tolerant species.
The trajectories of proportions among seed mass classes besides revealed the enhanced growth of small seeded, large dispersive, species and therefore the importance of species reproductive strategy which further supported the role of species dispersion and demographic strategy for the post-disturbance dynamics, as suggested by the IDH [@TerSteege2001;@Flores2006;@Haddad2008].

## On the recovery of disturbed communities

The divergence from the initial state however was very variable among plots, specifically after intense disturbance, which meant a strong dependency of plots recovery to initial composition or to abiotic conditions.
All plots, though, followed the same unimodal trajectory which confirmed both the resilience of forests taxonomic and functional composition and the convergence of communities towards initial states [@Hubbell1999;@Molino2001;@Baraloto2012a].
Functional traits are the most direct link between biodiversity and ecosystem functioning [@Diaz2005] the recovery of communities functional characteristics would go along the recovery of ecosystems processes although taxonomic composition remain altered [@Guariguata2001].
#The comparison of observed trajectories to a null model with random traits set shuffling showed an oriented functional shift after disturbance which suggest additional biotic and abiotic filters for the 10 to 15 years after disturbance.
#This means a shift from exclusive competition favoring competitive species before disturbance to environmental filtering selecting acquisitive and disturbance resistant species.

### 2018-04-16 WholePlotTrajectories.tex
Discussion

\subsection{Validation of the IDH}\label{validation-of-the-idh}

The enhanced communities evenness and functional richness after
disturbance confirmed the IDH for all treatment intensity. The
divergence from the initial state however was very variable among plots,
specifically after intense disturbance, which meant a strong dependency
of plots recovery to initial composition or to abiotic conditions. All
plots, though, followed the same unimodal trajectory which confirmed
both the resilience of forests taxonomic and functional composition and
the convergence of communities towards initial states
\citep{Hubbell1999, Molino2001, Baraloto2012a}. Functional traits are
the most direct link between biodiversity and ecosystem functioning
\citep{Diaz2005} the recovery of communities functional characteristics
would go along the recovery of ecosystems processes although taxonomic
composition remain altered \citep{Guariguata2001}. Disturbance proved to
favor species with acquisitive strategies, displaying efficient resource
acquisition (high SLA) but short-lived tissues (low wood density, leaf
toughness and thickness) and with a dispresive strategy (low seed mass)
\citep{Westoby1998, Wright2004, Reich2014}. \#The comparison of observed
trajectories to a null model with random traits set shuffling showed an
oriented functional shift after disturbance which suggest additional
biotic and abiotic filters for the 10 to 15 years after disturbance.
\#This means a shift from exclusive competition favoring competitive
species before disturbance to environmental filtering selecting
acquisitive and disturbance resistant species. In particular, the
evolution seed mass classes proportions revealed a clear increase in
small seeded, therefore large dispersive, species
\citep{TerSteege2001, Flores2006, Haddad2008}.

>>>>

\subsection{A validation of the intermediate disturbance
hypothesis}\label{a-validation-of-the-intermediate-disturbance-hypothesis}

The monitoring 30 years of disturbed forest communities confirmed the
limited direct impact of low disturbance on species richness as it was
found for impact survey of selective logging
(\citet{Cannon1998};\citet{Baraloto2010}). Only the most disturbed plots
had not fully achieved the ongoing recovery of genus richness, although
all had already reached equivalent levels as those of some control
plots. Such long time richness recovery had already been observed in
several logging experiments, which also highlighted the role of random
disturbance damage in the richness decrease
(\citet{deAvila2015};\citet{Hu2018}).

The evenness of all disturbed plots and the richness of low disturbance
plots on the other hand mainly followed an asymptotic growth with a
sharp increase until 15 years after disturbance. Such increased evenness
is explained by a much more homogeneous species distribution after
disturbance. Along with the IDH, disturbance then enhanced the growth
and survival of previously infrequent species, reorganizing the typical
high dominance structure of hyperdiverse mature forests. This increase
in taxonomic diversity was accompanied by an increased taxonomic
dissimilarity to initial state along time, translating a a taxonomic
turnover benefiting to pioneers and light demanding species, as
illustrated by the communities functional shifts towards
resource-acquisitive strategies (sharp increase in the SLA, leaf
thickness and bark thickness and decrease in wood density, leaf
toughness and maximum height)
\citep{Westoby1998, Wright2004, Reich2014}. In line with the IDH, the
post-disturbance changes in abiotic environment and competitive pressure
then well favored pioneers outcompeting other species when abiotic
resources are abundant but otherwise excluded in mature forests by
long-lived, resistant and shade tolerant species. The trajectories of
proportions among seed mass classes besides revealed the enhanced growth
of small seeded, large dispersive, species and therefore the importance
of species reproductive strategy which further supported the role of
species dispersion and demographic strategy for the post-disturbance
dynamics, as suggested by the IDH
\citep{TerSteege2001, Flores2006, Haddad2008}.

\subsection{On the recovery of disturbed
communities}\label{on-the-recovery-of-disturbed-communities}

The divergence from the initial state however was very variable among
plots, specifically after intense disturbance, which meant a strong
dependency of plots recovery to initial composition or to abiotic
conditions. All plots, though, followed the same unimodal trajectory
which confirmed both the resilience of forests taxonomic and functional
composition and the convergence of communities towards initial states
\citep{Hubbell1999, Molino2001, Baraloto2012a}. Functional traits are
the most direct link between biodiversity and ecosystem functioning
\citep{Diaz2005} the recovery of communities functional characteristics
would go along the recovery of ecosystems processes although taxonomic
composition remain altered \citep{Guariguata2001}. \#The comparison of
observed trajectories to a null model with random traits set shuffling
showed an oriented functional shift after disturbance which suggest
additional biotic and abiotic filters for the 10 to 15 years after
disturbance. \#This means a shift from exclusive competition favoring
competitive species before disturbance to environmental filtering
selecting acquisitive and disturbance resistant species.
